### 2019-07-17 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

---
title: "Transferability of clinical prediction models for early trauma care"
author: 
- Martin Henriksson
- Dell Saulnier
- Johanna Berg
- Martin Gerdin Wärnberg
date: "July 03, 2019"
documentclass: "article"
output:
  word_document:
    reference_docx: transferstyle.docx
header-includes:
   - \usepackage{hyperref}
bibliography: library.bib
csl: bmcemerg.csl
---
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

## LOAD FUNCTIONS, ENVIROMENT AND TABLES
  ## Source functions
  source("functions/WorkTableCreator.R")
  ## Retrive original results
  results <- readRDS("output/original.results.Rds")
  ## Retrive data frame with compiled results
  result.tables.data.frame <- readRDS("result.tables.data.frame.Rds")
  ## Create table from result.tables.data.frame for ease of data retrival in document
  results.all.data <- WorkTableCreator(result.tables.data.frame)

## RESULTS
  ## Total patients and NA
  total.number.of.patients <- nrow(results$data.sets.before.imputations$multi.centre.vs.single.centre$multi.centre)
  total.number.of.na <- results$NA.info.sample$multi.centre.vs.single.centre$multi.centre$total.number.of.na
  ## Sample NA
  na.vector <- unlist(results$NA.info.sample)
  names(na.vector) <- c("high volume", "high volume", "low volume", "low volume", "metropolitan", "metropolitan", "non metropolitan", "non metropolitan", "multi centre", "multi centre", "single centre", "single centre")
  highest.percent.missing.number <- max(na.vector[na.vector<100])
  highest.percent.missing.name <- names((na.vector[na.vector == highest.percent.missing.number]))
  ## Variable NA
  number.variable.na <- max(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na)
  percent.variable.na <- results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$percentage[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na]
  name.variable.na <- row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na,])
  gcs.na <- (results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre) == "ed_gcs_sum",])
  sbp.na <- (results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre) == "ed_sbp_value",])
  ## Development and validation
  ## Imputation data
  high.low.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "high.volume"]
  metro.non.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "metropolitan"]
  multi.single.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "multi.centre"]
  max.imp <- max(results$original.stats$Total.imputations)
  ## Validation data
  sorted.validation.mistriage <- sort(results.all.data[row.names(results.all.data) == "Validation.mistriage",], decreasing = FALSE)
  lowest.val.mistriage <- sorted.validation.mistriage[1]
  val.undertriage <- results.all.data["Validation.undertriage", colnames(lowest.val.mistriage)]
  val.overtriage <- results.all.data["Validation.overtriage", colnames(lowest.val.mistriage)]
  highest.val.mistriage <- sorted.validation.mistriage[6]
  ## Comparison
  ## Transfer data
  sorted.transfer.mistriage <- sort(results.all.data[row.names(results.all.data) == "Transfer.mistriage",], decreasing = FALSE)
  lowest.tran.mistriage <- sorted.transfer.mistriage[1]
  highest.tran.mistriage <- sorted.transfer.mistriage[6]
  ## Comparison data
  sorted.comp.mistriage <- sort(results.all.data[row.names(results.all.data) == "Transferred.mistriage.minus.buddy.local.mistriage",], decreasing = FALSE)
  highest.comp.mistriage <- sorted.comp.mistriage[6]
  worst.transfer <- results$original.stats[results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median == max(results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median),]

## DISCUSSION
  ## First paragraph
  UI.interval <- sub("\\).*", "", sub(".*\\(", "", format(highest.comp.mistriage)))
  UI.top.bot.list <- strsplit(UI.interval, split = " - ")
  UI.top.bot <- unlist(UI.top.bot.list)

## TABLE SETUP
  # Table 1 = Sample characteristics table
  high.low <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
  names(high.low) <- c("characteristic", "level", "high volume", "low volume")
  met.nonmet <- results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan
  names(met.nonmet) <- c("characteristic", "level", "metropolitan", "non metropolitan")
  multi.single <- results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre
  names(multi.single) <- c("characteristic", "level", "multi centre", "single centre")
   # Create complete table one (test)
   complete.table.one <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$metropolitan)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$non.metropolitan)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$multi.centre)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$single.centre.11001)
   names(complete.table.one) <- c("characteristic", "level", "high volume", "low volume", "metropolitan", "non metropolitan", "multi centre", "single centre")
   complete.table.one$characteristic <- c("n", "Age", "Gender", "", "GCS", "SBP", "RR", "30DS", "", "ISS", "NISS", "ISS>15", "")
  # Table 2 = result.tables, cropped to only show validation
  table.2 <- results.all.data[1:3,]
  rownames(table.2) <- c("Mistriage", "Undertriage", "Overtriage")
   # Create complete table two (test) 
   complete.table.two <- results.all.data[1:3,]
   rownames(complete.table.two) <- c("mistriage", "undertriage", "overtriage")
  # Table 3 = Cropped to only show transfer
  table.3 <- results.all.data[4:6,]
  rownames(table.3) <- c("Mistriage", "Undertriage", "Overtriage")
  # Table 4 = Cropped to only show transfer differance
  table.4 <- results.all.data[7:9,]
  rownames(table.4) <- c("Mistriage", "Undertriage", "Overtriage")
```
---
abstract: |

  **Background:** Trauma results in approximately 3,8 million deaths annually. Clinical
  prediction models are often used for triage in early trauma care. Previous research has
  indicated transferring these models between different settings might adversely affect 
  model performance.

  **Methods:** Patients age 15 or above registered between the years 2011 and
  2016 in the Swedish national trauma registry SweTrau were included. Three sets
  of data were created: High and low volume centres, metropolitan and
  non-metropolitan centres, and multi and single centres. Clinical
  prediction models were developed using logistic regression in each data set 
  and transferred within sets. Model performance was evaluated in all data
  sets using mistriage rate, undertriage rate and overtriage rate. Multiple
  imputation using chained equation was used to handle missing data. Model
  performance was reported as medians with 95% uncertainty interval across
  imputed data sets.

  **Results:** `r total.number.of.patients ` patients were included. Model
  validation mistriage rate ranged from `r lowest.val.mistriage` 
  to `r highest.val.mistriage`, and transferred model
  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
  across all samples. Performance after transfer when compared to receiving sample validation 
  performance resulted in subtantial uncertainty in all samples. Worst performance after 
  transfer was found in the `r colnames(highest.comp.mistriage)` 
  sample with a mistriage difference of `r highest.comp.mistriage`.

  **Conclusions:** In all samples model transfer was associated with high 
  levels of uncertainty. Notably, transfer from the `r colnames(highest.comp.mistriage)` 
  sample to the other sample in the data set lead to an increased mistriage
  rate, primarily due to increased overtriage.

  **Keywords:** Trauma, Clinical prediction model, Transferability, Mistriage, Undertriage, Overtriage 
---

# Introduction
Trauma results in approximately 3,8 million deaths annually across the globe, 
and road traffic injuries are among the top 10 leading causes of death [@RN1; @RN21]. 
Each year 9% of the world’s deaths are because of trauma, and
predictions indicate that this number is likely to increase [@RN42].

In a typical, high resource setting, the initial trauma management is
performed on the scene by Emergency Medical Services
(EMS). Patient data and vital signs are transferred to the receiving hospital. 
This information is then evaluated using a system to determine
the level of trauma, prepare adequate resources [@RN43] and dictate if a full
or limited trauma team is activated [@RN44].

Systems to determine the level of trauma can be based on clinical prediction
models. Models differ in quality and characteristics, but generally perform
well predicting survival [@RN45]. Many of these models are developed in a single
standardized setting such as a major trauma centre but are then implemented in 
different settings, such as non-trauma centres.

What is not fully understood is how this transfer affects model
performance. Previous research has shown that model performance in terms of
calibration can be adversely affected [@RN27]. However, the aforementioned
research studied model transfers between substantially different settings and
did not assess more clinically relevant performance measures, such as
misclassification.

In trauma, misclassification is often referred to as mistriage. Triage refers to
the classification of severity as minor or major trauma. Mistriage can be
subdivided into overtriage; incorrect classification of a patient with minor
trauma as major trauma, or undertriage; incorrect classification of a patient
with major trauma as minor trauma. Both are detrimental in terms of patient care
and distribution of resources.

Thus, the effect of model transfers between contexts within the setting of a
single health care system, as well as the effect of such transfers on mistriage
have not been studied and represent substantial knowledge gaps. Therefore, the
aim of this study is to assess how transfers of clinical prediction models for
early trauma care between different contexts within a single health system
affect mistriage rates.

# Materials and methods
## Study design
A registry-based cohort study was conducted. Sweden has a nationally encompassing
trauma register called SweTrau. SweTrau data was used to create
clinical prediction models, which were transferred between different contexts to
study the effects on mistriage. The study and analysis plans were made
publically available before the research was undertaken [@RN59].

## Setting
In Sweden, trauma is the most common cause of death before 44 years of age. Up 
to 10% of all major trauma patients suffer some form of disability. 
The societal effects of trauma are also substantial, with the loss
of work days comparable to those lost from cardiovascular and malignant diseases
combined [@RN23].

The trauma care organization in Sweden is usually clearly defined in major
metropolitan areas: trauma patients are transported directly to a predesignated
hospital with specific trauma competency. In large, more rural parts of the
country, trauma patients are instead transported to the nearest hospital for
stabilization. Once stabilized, they may then be transported to hospitals with
more advanced trauma care capacity [@RN34]. Most hospitals in Sweden use some
form of system to categorize the level of trauma. These systems are usually
based on patient vital signs and injury mechanism, as reported by the EMS, or as
registered in the emergency department [@RN90].

Today 95,5% (52 of 55) of Swedish hospitals record trauma cases in SweTrau. 
Currently the registry includes 55 000 cases, and the board encourages its 
use in both academic research and local quality improvement initiatives [@RN28].

## Participants 
The eligibility criteria were patients registered in SweTrau between 2011 and 2016, 
age 15 or above. This age was decided as the study aims to study adult trauma and not
paediatric trauma which differs in physiology, triage and initial care
[@RN46]. Also, the Swedish guidelines for trauma activation define children as age <15
[@RN90]. Patient age was obtained from the SweTrau register.

## Variables
### Model predictors
The clinical prediction models used the predictors systolic blood pressure
(SBP), respiratory rate (RR) and Glasgow coma scale (GCS), all on arrival to
hospital. The rationale for these three predictors is that they are
part of many established clinical prediction models for early trauma care, such
as the Revised Trauma Score [@RN35].

### Model outcome
The outcome used to develop the clinical prediction models was all cause
mortality within 30 days of trauma.

### Participant characteristics
To describe the patient cohort we reported age, sex, Injury Severity Score (ISS)
and New ISS (NISS).

### Study outcome
ISS > 15 was used as the gold standard to define trauma severity as major
trauma, and hence patients with ISS ≤ 15 were considered minor trauma
[@RN36]. Overtriage was defined as the event when a clinical prediction model
classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event
when a clinical prediction model classified a patient with ISS > 15 as minor
trauma. Overtriage rate was defined as the number of overtriaged patients
divided by all patients. Undertriage rate was defined as the number of
undertriaged patients divided by all patients. Mistriage rate was defined as the
sum of the over- and undertriage rates.

## Data sources and measurements
Model predictors, outcome, participant characteristics and study outcome as
outlined above were all obtained from SweTrau. The method of measurement for the
model predictors is not specified in the registry entries. In Swedish emergency rooms,
patient parameters are usually obtained by a registered nurse or assistant nurse
and are assumed to be accurate.  Whether the patient is dead or alive 30 days
after trauma is manually entered into the registry locally by each respective
hospital. NISS and ISS are calculated by hospital personal
based on the patient’s injuries.

## Bias
Data analysis was conducted according to a prearranged analysis plan. The
analysis plan and statistical analysis code was finalised using simulated
data. These efforts were taken to avoid confirmation bias. The analysis plan was
reviewed by an experienced statistician and programmer prior to implementation. 
Neither outcome nor variable were blinded when conducting data analysis, which 
made a structured objective approach important.

## Study size
All patients matching eligibility criteria listed above. Four data sets were
used to study the transfer of clinical prediction models, each data set
representing a different setting. Data set, and sample size considerations are
outlined below.

## Quantitative variables
SBP and RR was modelled using restricted cubic splines with four knots whenever
possible placed at equally spaced percentiles, and GCS as a continuous linear
term. When describing the participant characteristics quantitative variables
were presented as continuous. ISS was presented as dichotomous using ISS > 15 as
the cutoff.

## Statistical methods
### Data sets
The complete SweTrau cohort was split into four sets of data, each representing a 
plausible clinical setting for model transfer. These data sets were further divided 
into samples, this process is outlined below.

### High and low volume centres
Based on number of patients received and registred in SweTrau for each centre, two 
samples were derived from this data set. High volume centres were those with in the 
top quartile of number of patients registred. The rest were low volume centres.

### Metropolitan and non-metropolitan centres
This data set was also split into two samples. The metropolitan sample consisted
of greater Stockholm, greater Gothenburg and greater Malmö, as defined by
statistics Sweden. The other sample was patients from non-metropolitan areas.

### Multi and single centre data
In this data set multiple samples were created. Each centre with large enough
sample size (outlined below) to develop and validate a model constituted their own 
sample. The multi-centre sample consisted of the combined data from all centres.

### Individual centres
This data set was also split into multiple samples. Each centre with large
enough sample size to develop and validate a model constituted its own
sample. This data set was removed in the final version of the study due to a
lack of centres with a large enough sample size.

### Development and validation sample
Each data set included at least two samples. The samples were then split into
two subsamples using a temporal split based on the date of trauma. The earlier
subsample was the development sample, and the later subsample the validation
sample. The development sample contained 70 events (events being patients who
died within 30 days of the trauma) and all non-events (non-events being patient
survival 30 days past the trauma) during the same time. The rationale for
including 70 events is that at least 10 events per free parameter were needed in
the logistic regression to obtain stable coefficient estimates [@RN30]. The
validation sample instead contained 100 events and at least 100 non-events,
which was suggested as the minimum number by Vergouwe in 2005 for external
validation samples [@RN37]. See figure 1 for example using the high- and low
volume centre data set. The minimum sample size of development and
validation samples were 140 and 200 respectively. We performed analyses only on
data sets for which all samples included at least the minimum sample size.

![](fig1.jpg)  
**Fig. 1: High and low volume centre data set.** Illustrates creation of development 
and validation sample in the high and low volume data set. Initial split based on
number of patients. Temporal split made using date of traumatic event.

### Sequence of analysis
The programming language R was used for all analyses [@RN31], and all code has
been made publically available [@RN59]. Analyses were performed in the sequence
of model development, model validation and finally model comparison. These steps
were repeated in each data set. Below we used the transfer of a model from a
high-volume sample to a low volume sample as an example to describe the complete
procedure.

### Model development
In the model development step a clinical prediction model was developed in the
high-volume centre development sample. The model was developed using logistic
regression. The dependent variable was all cause mortality within 30 days of 
trauma and independent variables SBP, RR, and GCS modelled as previously described. 
To avoid overfitting the model a bootstrap procedure was used to estimate a linear 
shrinkage factor that was then applied to the model coefficients [@RN61]. The shrunk 
model was then used to estimate the probability of all cause 30-day mortality in the 
development sample. A gridsearch was performed across estimated probabilities in 
the development sample to identify the cut-off that optimised overtriage keeping 
undertriage at less than 5% [@RN39]. This cut-off was then used to classify patients 
as major or minor trauma.

### Model validation
In the model validation step, the model performance was assessed in the
high-volume centre validation sample and in the low volume centre validation
data. First the model was used to estimate the probability of all cause 30-day
mortality in each sample. Then the probability cutoff identified in the
development sample was applied to the validation samples, patients were
classified as major or minor trauma, and model performance was estimated.

### Model comparison
Finally, in the model comparison step, the difference in model performance
between the high and low volume centre validation samples was
calculated. Empirical bootstrap was used to estimate 95% confidence intervals
(CI) around performance and differences in performance estimates. Both bootstrap
procedures used 1000 bootstrap samples drawn with replacement of the same size
as the original samples. The three steps of model development, model validation,
and model comparison were repeated in all four sets of data.

### Performance measures
Model performance was assessed in terms of over-, under-, and mistriage rates as
defined above.

### Missing data
We used multiple imputation using chained equations (mice), to handle missing data [@RN41]. 
The number of imputations created
for each data set was equal to the highest percentage of missing data in that
data set. Quantitative variables were imputed using predictive mean matching and
qualitative variables were imputed using logistic regression. SBP and RR was
transformed as restricted cubic splines before imputation. All analyses outlined 
above were then conducted separately in
each imputed dataset. We present the combined results as the median point
estimate across imputations along with an empirical bootstrap of the 25th and
75th percentiles across imputations, i.e. the lower bound of the presented
interval is the lower bound of a 95% CI of the 25th percentile and the upper
bound is the upper bound of a 95% CI of the 75th percentile. This combined CI
was referred to as an Uncertainty Interval (UI) and was used to express the
added uncertainty associated with the imputation procedure to handle missing
data, as such it is more conservative than a standard 95% CI.

## Ethical considerations
This study is approved by the regional ethics review board in Stockholm,
Sweden. Ethical review numbers are 2015/426-31 and 2016/461-32. All patients
included recived letters detainling SweTrau inclusion, and the possibility of
their data being used for scientific purposes. It was assumed that the patients
who consented to research want to contribute to this improvement of care, and as
thus by performing this study the authors hope to honour their wishes.

# Results
We analysed data from `r total.number.of.patients ` trauma patients to
investigate the effects of transfer on clinical prediction models (Table 1). The
total number of missing observations across all variables was `r total.number.of.na ` 
in the entire study cohort. The sample with the highest
percentage of missing observations was the `r highest.percent.missing.name `
sample with `r highest.percent.missing.number `% incomplete observations. The
variable with the highest number of missing values was `r name.variable.na `,
with `r number.variable.na ` missing values, which constituted `r round(percent.variable.na) `
% of the total values for this variable. The percentages 
of missing values for the other model predictors, GCS and SBP was `r round(gcs.na$percentage)`% 
and `r round(sbp.na$percentage)`% respectively.

**Table 1. Sample characteristics of multiple imputed data.** Data is presented as medians with interquartile range (IQR) or counts with % as applicable.
```{r, echo = FALSE, results = "asis"} 
## Table 1
options(knitr.kable.NA = '')
kable(high.low, format = "markdown")
kable(met.nonmet, format = "markdown")
kable(multi.single, format = "markdown")
# kable(complete.table.one, format = "markdown")
``` 
GCS, glasgow coma scale; SBP, systolic blood pressure; RR, respiratory rate; 30DS, 30 day survival; ISS, injury severity score; NISS, new injury severity score

## Development and validation
During model development, the number of imputations used for each data set was
`r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high and
low volume data set, the metropolitan and non-metropolitan data set and the
multi and single centre data set respectively.

Table 2 shows model validation performance. The model with the lowest validation
mistriage rate was the `r colnames(lowest.val.mistriage)` model with a median
mistriage rate of `r lowest.val.mistriage`. Performance in terms of
undertriage and overtriage rate for the same model was `r val.undertriage`, and
`r val.overtriage` respectively. Worst validation performance i.e. highest
mistriage rate was found in the `r colnames(highest.val.mistriage)` sample, with
a model median mistriage rate of `r highest.val.mistriage`.

**Table 2. Model validation performance, i.e. performance when applied to the sample in which the model was created.** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).
```{r, echo = FALSE, results = "asis"} 
## Table 2
kable(table.2[,1:2], format = "markdown")
kable(table.2[,3:4], format = "markdown")
kable(table.2[,5:6], format = "markdown")
# kable(complete.table.two, format = "markdown")
```

## Comparison
Model performance after transfer was determined for each model being transferred
to the other sample in the data set, in all data sets (Table 3). When
transferred, the model with the lowest mistriage rate was the `r colnames(lowest.tran.mistriage)` 
model with a median mistriage rate of `r lowest.tran.mistriage`. The model with 
the highest mistriage rate after transfer was the `r colnames(highest.tran.mistriage)` 
with a median mistriage rate of `r highest.tran.mistriage`.

**Table 3. Model transfer performance, i.e. performance when transferred to the other sample in the same data set. ** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).  
```{r, echo = FALSE, results = "asis"} 
kable(table.3[,1:2], format = "markdown")
kable(table.3[,3:4], format = "markdown")
kable(table.3[,5:6], format = "markdown")
```

Table 4 shows the transferred model performance when compared to the validation
performance in the sample reciving the transferred model.
The highest mistriage rate when compared to the reciveing sample validation
model performance was found in the `r colnames(highest.comp.mistriage)` model,
with a median mistriage rate difference of `r highest.comp.mistriage`, meaning
this model performed worse when transferred to the other sample in the same data
set. In clinical terms, this model transfer means that among 100 trauma
patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` 
more patients would be wrongly classified as major or minor trauma.

**Table 4. Comparison mistriage, i.e. sample validation performance subtracted from transferred performance.** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).  
```{r, echo = FALSE, results = "asis"} 
kable(table.4[,1:2], format = "markdown")
kable(table.4[,3:4], format = "markdown")
kable(table.4[,5:6], format = "markdown")
```

# Discussion
This study aimed to assess how transfers of clinical prediction models for early
trauma care between different contexts within a single health system affect
mistriage rates. In all data sets, model transfer between samples resulted in
substantial uncertainty in its effect on model performance. When comparing
transferred performance to the reciveing sample validation performance the
resulting differance also showed considerable uncertainty.

The most notable effects on model performance following model transfer were seen
after transferring the `r colnames(highest.comp.mistriage)` model. This transfer
resulted in an increase of the mistriage rate by 
`r substr(format(highest.comp.mistriage), 1, 5)`.  Mainly contributing to this was
a marked increase in overtriage.  For this transfer the 95% UI ranged from 
`r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty and
that our findings are compatible with a `r UI.top.bot[1]` decrease to a 
`r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]`
would signifiy an improvement in model performance. The higher 95% UI of 
`r UI.top.bot[2]` would mean a marked increase in model mistriage, compared to the
point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.

The transfer of clinical prediction models in trauma care has not previously
been studied extensively. In 2016, Gerdin et al. found model transfer to
adversely affect model performance in terms of calibration, but that this could
be improved by updating the model [@RN27]. This study did not examine model
performance in terms of calibration, neither were the effects of updating the
model explored.

External validation as the evaluation of a clinical prediction model in a
setting in which it was not originally developed has been studied more
extensively. Studies using simulated clinical prediction models based on
different predictors showed a decline in model performance when being externally
validated [@RN60; @RN61]. A study from 2018 examining the external validity of
prediction models for coronary artery disease (CAD) [@RN62], showed models
significantly underestimating the probability of CAD. The combined results of
this study, the studies on model transfer and the studies on external validation
suggests caution when transferring models.

## Strengths and limitations
### Strengths
The main strength of this study is that its design realistically reflects
potential model transfers: clinical prediction models for trauma are usually
developed in large metropolitan centres and then transferred as clinical
recommendations to minor non-metropolitan centres. The results should therefore
be of practical importance.

### Limitations
One limitation is that originally all single centres with a valid number of 
events were to constitute a data set, but only one centre provided a sufficient 
number of events. Leading to the loss of the data set: Individual centres. 
This data set might have provided results that could reflect realistic 
situations as major trauma centres regularly exchange data.

Furthermore, this study used GCS, RR and SBP as predictors for 30-day
mortality. Additional predictors such as age, mechanism
of injury, or newer predictors as shock index might increase model
performance. Furthermore, we used 30-day mortality as outcome, recognizing
that late mortality, time in hospital, functional impairment, and morbidity are
also important outcome factors in the context of trauma care.

The predetermined analysis plan was to be followed in a step-by-step fashion, to
avoiding confirmation bias. Programming challenges and late-time code inclusions
required a flexible approach to this analysis plan, sometimes with revisions of
previous steps. This is not optimal, and efforts were made to not deviate in any
major way. Using GitHub, all code was made public for inspection and to ensure
reproducibility [@RN59].

## Clinical applications
This study shows that mistriage can potentially increase following model
transfer, and more prominently in certain cases. This supports the view that
model transfers should be done with caution, as performance might
deteriorate.

Recognizing these risks and updating models accordingly should lead
to increased accuracy during trauma patient triage. Potentially saving lives by
minimizing undertriage, and likely save resources by minimizing overtriage. With
registries such as SweTrau, analysis of model transfer could be performed
continuously to optimize models currently in use and advise which model
transfers might be detrimental (or potentially beneficial).

## Conclusion
Model transfer resulted in large and varying levels of uncertainty in its effect
on mistriage. Therefore model transfer can be unpredictable, potentially leading 
to increased mistriage, but in some cases also to a decrease in mistriage. 
Noticeably, transfer of the `r colnames(highest.comp.mistriage)` model lead to an 
increased mistriage of `r highest.comp.mistriage`. This observed effect was also 
associated with considerable uncertainty. Recognizing the limitations of this study, 
the authors believe further studies are warranted dueto the potential economic 
costs and patient consequences of poor model transfers.

# References